\documentclass[conference]{IEEEtran}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{enumitem}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,urlcolor=blue}

\title{Guardrails-as-a-Service: Hybrid Deterministic and Semantic Policy Evaluation with Tamper-Evident Auditability}

\author{
\IEEEauthorblockN{(Author Name Redacted)}
\IEEEauthorblockA{Email: contact@example.com}
}

\begin{document}
\maketitle

\begin{abstract}
We present the architecture and planned evaluation of a cloud-agnostic Guardrails-as-a-Service platform that integrates deterministic rule evaluation with semantic vector retrieval to enforce complex governance policies for AI-enabled systems. The design emphasizes scalability via stateless microservices and asynchronous event flow, extensibility through a handler abstraction, and audit integrity using hash-chained decision logs. We define measurable quality attributes (latency, throughput, extensibility effort, audit lag) and outline empirical validation strategies supporting operational assurance and trust.
\end{abstract}

\begin{IEEEkeywords}
Policy enforcement, microservices, vector similarity, audit logging, hash chaining, observability, cloud governance.
\end{IEEEkeywords}

\section{Introduction}
Dynamic policy enforcement in modern distributed and AI-centric systems requires synthesizing explicit rules with semantic context. Traditional engines (e.g., Open Policy Agent (OPA) \cite{opa}) optimize declarative logic but lack native semantic retrieval. Cloud-specific governance (e.g., AWS Control Tower guardrails \cite{awsguardrails}) constrains portability. We address these gaps with a hybrid evaluation plane coupling rule logic and embedding similarity, backed by an integrity-focused audit pipeline.

\section{Contributions}
The work provides: (1) A hybrid evaluation architecture combining deterministic rule filtering and vector similarity; (2) An extensible handler interface enabling semantic and emerging policy types with low change cost; (3) A hash-linked audit log for tamper-evidence without blockchain overhead; (4) A metrics-driven evaluation plan aligning architectural tactics with verifiable quality attributes.

\section{System Architecture}
Fig.~\ref{fig:arch} conceptually partitions concerns: policy lifecycle management, evaluation, event streaming, caching, and observability. Decoupling via a message broker enables back-pressure handling and horizontal scale.

\noindent Core components:
\begin{itemize}[leftmargin=*]
\item API Gateway: Authentication, authorization, rate limiting.
\item Policy Service: CRUD, versioning, vector indexing (pgvector \cite{pgvector}).
\item Evaluation Plane: Cache-first retrieval, rule + semantic evaluation, decision event emission.
\item Audit Logger: Consumes decision events; persists hash-chained records.
\item Redis: Low-latency hot-set and embedding cache.
\item Broker: Decouples evaluation from persistence; supports resilience tests.
\item Observability Stack: Metrics (Prometheus \cite{prometheus}), logs, planned traces (OpenTelemetry \cite{otel}).
\end{itemize}

\section{Design Methodology}
Microservice decomposition isolates policy mutation from evaluation throughput. Event-driven processing reduces synchronous coupling. Vector similarity augments metadata filtering, allowing semantically proximate policy selection. Hash chaining supplies integrity assurance with linear verification.

\section{Quality Attributes}
\subsection{Scalability}
Tactics: Stateless pods, horizontal autoscaling, Redis caching, asynchronous buffering. Metrics: P95 latency $<150$\,ms under 2000 requests/sec; cache hit ratio correlated with latency reduction.

\subsection{Extensibility}
A \texttt{PolicyHandler} interface:
\begin{lstlisting}[language=TypeScript,basicstyle=\ttfamily\small]
export interface PolicyHandler {
  supports(policyType: string): boolean;
  evaluate(context: any, policy: Policy): Promise<Decision>;
}
\end{lstlisting}
New handlers (e.g., advanced semantic constraint) integrate via dependency injection without core modification. Verification: Diff locality; unchanged regression suite; cyclomatic complexity < 10.

\subsection{Observability \& Auditability}
Correlation IDs propagate through logs and metrics. Audit integrity uses chained hashes:
\begin{lstlisting}[language=Python,basicstyle=\ttfamily\small]
prev_hash = "GENESIS"
for event in decision_events:
    serialized = canonical_json(event)
    curr_hash = sha256(prev_hash + serialized)
    store({...event, prev_hash, curr_hash})
    prev_hash = curr_hash
\end{lstlisting}
Periodic verification recomputes sequence and alerts on mismatch; target audit lag < 5\,s at sustained load.

\subsection{Maintainability}
Modular boundaries and standardized evaluation signatures reduce change effort. KPIs: Mean Time To Change for routine policy extension < 0.5 developer-day.

\section{Implementation Overview}
Policies stored in PostgreSQL with pgvector embeddings; evaluation pipeline performs:
(1) Cache lookup;
(2) Vector similarity search (threshold-based);
(3) Rule evaluation;
(4) Decision event emission;
(5) Audit service hash-link persistence.

\section{Evaluation Plan}
Phases: (1) Functional tests (unit, integration); (2) Load tests (k6 ramp to 2000 req/sec); (3) Resilience (broker partition throttling). Metrics scraped every 5\,s; Grafana dashboards visualize throughput, latency, audit lag. Success: No message loss; hash chain validates end-to-end; latency targets met.

\section{Security Considerations}
Current: JWT-based client auth; least-privilege DB roles; rate limiting. Planned: mTLS inter-service; policy signature verification for provenance.

\section{Related Work}
OPA \cite{opa} offers strong declarative evaluation but lacks semantic retrieval integration and native hash-chained audit logging. AWS Control Tower guardrails \cite{awsguardrails} provide opinionated cloud governance with limited cross-cloud extensibility. Our approach differentiates through hybrid retrieval, event-driven decoupling, and cryptographic audit integrity.

\section{Conclusion}
The proposed Guardrails-as-a-Service platform aligns architectural tactics with measurable quality attributes. Hybrid policy evaluation and hash-linked auditing together enhance governance fidelity and trust. Future work targets production hardening, expanded semantic handlers, and full distributed tracing.

\section*{Acknowledgment}
Omitted for anonymity.

\begin{thebibliography}{99}
\bibitem{opa} Open Policy Agent Documentation. [Online]. Available: \url{https://www.openpolicyagent.org}
\bibitem{awsguardrails} AWS Control Tower Guardrails. [Online]. Available: \url{https://docs.aws.amazon.com/controltower}
\bibitem{prometheus} Prometheus Documentation. [Online]. Available: \url{https://prometheus.io}
\bibitem{pgvector} pgvector Extension Documentation. [Online]. Available: \url{https://github.com/pgvector/pgvector}
\bibitem{otel} OpenTelemetry Specification. [Online]. Available: \url{https://opentelemetry.io}
\end{thebibliography}

\end{document}
